---
title: "Group Project 2"
subtitle: "Biology 368/664 Bucknell University"
output: html_notebook
author: Prof. Ken Field
date: 19 Feb 2019
---

## Objectives

1. Critically read a scientific paper and assess the evidence provided
2. Assess a published paper for issues with reproducibility and transparency
3. Identify questionable research practices and describe appropriate methods to prevent them
4. Complete a preregistration plan describing a confirmatory study following the principles of 

## Groups

We will work in pairs for this project, with each pair being different from the Project 1 groups.

```{r Load Libraries, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
if (!require("knitr")) install.packages("knitr"); library(knitr)
```

```{r echo=FALSE, results='asis'}
Project2 <- read.csv("../../Biol364_664roster.csv", stringsAsFactors=FALSE)
Project2$Group <- c("C",
                    "A",
                    "D",
                    "B",
                    "F",
                    "C",
                    "A",
                    "D",
                    "E",
                    "B",
                    "F",
                    "E")
table(Project2$Group)
kable(arrange(Project2, Group))
```

## Choosing Papers

Each group will choose a paper in lab on Tuesday, Feb 19. You may want to browse the papers available and discuss them with you group partner prior to lab.

```{r}
dir("papers", pattern = "pdf", full.names = TRUE, ignore.case = TRUE)
```

## Grading

Each student will be expected to complete the following tasks to earn 85% of the points available for this assignment (21/25).

- Read and discuss your paper with your group partner
  + Focus on the methods and results
  + For each conclusion evaluate the strength of the supporting evidence
- Complete the nr-reporting-summary.pdf checklist (open in Adobe Acrobat)
  + For any deficient items on the checklist, describe the information needed
  + Refer to nr-reporting-life-sciences-research.pdf for guidelines
- Determine if you could repeat the analysis that is most essential to conclusions of the paper
  + If data or methods are not available, look for follow-up studies that provide the required details
  + If follow-up studies are not available, describe the missing information needed to perform the analysis
- For each Questionable Research Practice discussed in Fraser *et al*:
  + Find evidence to exclude this QRP, or
  + Describe steps that should have been taken to prevent this QRP (see Forstmeier *et al*)
- Design a follow-up study to replicate the most crucial experiment
  + Complete the PreregChallengeTemplate.Rmd form for this replication experiment
- Present summary of reproducibility assessment in class

Each group should use an **Acknowledgements** section to document the participation of each member and the collaboration within and between groups.

Additional credit will be awarded for providing assistance to other groups and for completion of the analysis that supports the most essential conclusion.


