---
title: "Group Project 2"
subtitle: "Biology 368/664 Bucknell University"
output: html_notebook
author: Prof. Ken Field
date: 19 Feb 2019
---

## Objectives

1. Critically read a scientific paper and assess the evidence provided
2. Identify questionable research practices and describe appropriate methods to prevent them
3. Complete a preregistration plan describing a confirmatory study following the principles of 

## Groups

We will work in pairs for this project, with each pair being different from the Project 1 groups.

```{r Load Libraries, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
```

```{r echo=FALSE}
Biol364_664roster <- read.csv("../../Biol364_664roster.csv", stringsAsFactors=FALSE)
Project1 <- Biol364_664roster
Project2 <- Biol364_664roster
Project2$Group <- c("C",
                              "A",
                              "D",
                              "B",
                              "F",
                              "C",
                              "A",
                              "D",
                              "E",
                              "B",
                              "F",
                              "E")
Project2 %>%
  arrange(Group) %>% 
  tbl_df %>% 
  print(n=12)
```

## Grading

Each student will be expected to complete the following tasks to earn 85% of the points available for this assignment (21/25).

- Identify and obtain suitable dataset
- Create Github repository and use version control to collaborate on project
- Spend 4-6 hours preparing, coding, and testing tutorial
  + Data exploration
  + Data visualization
  + Hypothesis testing
- Present tutorial in class
- Provide public archive suitable for sharing to students/faculty

Each group should use an *Acknowledgements* section to document the participation of each member and the collaboration within and between groups.

Additional credit will be awarded for providing assistance to other groups and for the development of a tutorial that goes beyond the minimal expectations listed above.


