\documentclass[11pt,]{article}
\usepackage[margin=1in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{mathpazo}
\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center
\newcommand{\blankline}{\quad\pagebreak[2]}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}} 
\usepackage{longtable,booktabs}

\usepackage{parskip}
\usepackage{titlesec}
\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{6pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{6pt plus 2pt minus 2pt}

\titleformat*{\subsubsection}{\normalsize\itshape}

\usepackage{titling}
\setlength{\droptitle}{-.25cm}

%\setlength{\parindent}{0pt}
%\setlength{\parskip}{6pt plus 2pt minus 1pt}
%\setlength{\emergencystretch}{3em}  % prevent overfull lines 

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{lastpage}
\renewcommand{\headrulewidth}{0.3pt}
\renewcommand{\footrulewidth}{0.0pt} 
\lhead{}
\chead{}
\rhead{\footnotesize Biology 364/664: Analysis and Visualization of Biological Data -- Spring 2019}
\lfoot{}
\cfoot{\small \thepage/\pageref*{LastPage}}
\rfoot{}

\fancypagestyle{firststyle}
{
\renewcommand{\headrulewidth}{0pt}%
   \fancyhf{}
   \fancyfoot[C]{\small \thepage/\pageref*{LastPage}}
}

%\def\labelitemi{--}
%\usepackage{enumitem}
%\setitemize[0]{leftmargin=25pt}
%\setenumerate[0]{leftmargin=25pt}




\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
}
\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={ ()},
             pdfkeywords = {},  
            pdftitle={Biology 364/664: Analysis and Visualization of Biological Data},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls


\setcounter{secnumdepth}{0}





\usepackage{setspace}

\title{Biology 364/664: Analysis and Visualization of Biological Data}
\author{Prof.~Ken Field}
\date{Spring 2019}


\begin{document}  

		\maketitle
		
	
		\thispagestyle{firststyle}

%	\thispagestyle{empty}


	\noindent \begin{tabular*}{\textwidth}{ @{\extracolsep{\fill}} lr @{\extracolsep{\fill}}}


E-mail: \texttt{\href{mailto:kfield@bucknell.edu}{\nolinkurl{kfield@bucknell.edu}}} & Web: \href{http://\url{https://github.com/KField-Bucknell}}{\tt \url{https://github.com/KField-Bucknell}}\\
Office Hours: T 12:00-1:00 pm  &  Class Hours: TR 11:00-11:52 am, T 1:00-5:00 pm\\
Office: 208 Biology Building  & Class Room: TBA\\
	&  \\
	\hline
	\end{tabular*}
	
\vspace{2mm}
	


\hypertarget{course-description}{%
\section{Course Description}\label{course-description}}

Introduction to the process of data exploration and visualization using
state-of-the-art computational techniques. Using ``big data'' from
public archives or their own research projects, students will learn how
to rigorously analyze and visualize complex biological datasets. Lab
will include hands-on work with R and virtual reality. No programming
experience required.

\hypertarget{course-objectives}{%
\section{Course Objectives}\label{course-objectives}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Using R, students will analyze, visualize, and intepret real-world
  datasets.
\item
  Students will learn to apply rigorous standards to experimental design
  and interpretation.
\item
  Working as a team, students will complete novel projects utilizing
  whole-transcriptome or whole-genome datasets.
\item
  Students will present their final projects using complex
  multi-dimensional data visualizations.
\end{enumerate}

\hypertarget{required-readings}{%
\section{Required Readings}\label{required-readings}}

\hypertarget{course-policy}{%
\section{Course Policy}\label{course-policy}}

I will detail the policy for this course below. Basically, don't cheat
and try to learn stuff. Don't be that guy.

\hypertarget{grading-policy}{%
\subsection{Grading Policy}\label{grading-policy}}

\begin{itemize}
\item
  \textbf{20\%} of your grade will be determined by a midterm during
  normal class hours.
\item
  \textbf{20\%} of your grade will be determined by a term paper that
  documents your appreciation of Foghat's ``Slow Ride'', the most
  important song ever written. ``Slow Ride'' is what Mozart wishes
  \emph{Don Giovanni} could have been.
\item
  \textbf{10\%} of your grade will be determined by your attendance and
  participation in class. Generally, ask questions and answer them.
\item
  \textbf{20\%} of your grade will be determined by a 20-page term paper
  on when exactly ``The Love Boat'' jumped the proverbial shark. You
  will address whether this shark-jumping can be attributed to Ted
  McGinley, the introduction of Jill Whelan as ``Vicki'', or some other
  cause.
\item
  \textbf{30\%} of your grade will be determined by a final exam.
\end{itemize}

\hypertarget{attendance-policy}{%
\subsection{Attendance Policy}\label{attendance-policy}}

\hypertarget{e-mail-policy}{%
\subsection{E-mail Policy}\label{e-mail-policy}}

\hypertarget{make-up-exam-policy}{%
\subsection{Make-Up Exam Policy}\label{make-up-exam-policy}}

There are \textbf{NO} make-ups for missed exams. Don't bother asking.

\hypertarget{academic-dishonesty-policy}{%
\subsection{Academic Dishonesty
Policy}\label{academic-dishonesty-policy}}

Don't cheat. Don't be that guy. Yes, you. You know exactly what I'm
talking about too.

\hypertarget{disabilities-policy}{%
\subsection{Disabilities Policy}\label{disabilities-policy}}

Federal law mandates the provision of services at the university-level
to qualified students with disabilities. Make sure to include all that
relevant information here.

\newpage

\hypertarget{class-schedule}{%
\section{Class Schedule}\label{class-schedule}}

Students must read the following before Tuesday's class session.
Important: class readings are subject to change, contingent on
mitigating circumstances and the progress we make as a class. Students
are encouraged to attend lectures and check the course website for
updates.

\hypertarget{week-01-0114---0118-syllabus-day}{%
\subsection{Week 01, 01/14 - 01/18: Syllabus
Day}\label{week-01-0114---0118-syllabus-day}}

\emph{No class Thursday (Political scientists usually have a conference
to start the semester).}

Read \emph{all} associated documents on course website.

\begin{itemize}
\tightlist
\item
  \href{http://svmiller.com/blog/2014/09/taking-good-notes/}{Taking Good
  Notes}
\item
  \href{http://svmiller.com/blog/2015/06/dos-and-donts-of-writing-for-students/}{Dos
  and Dont's of Writing for Students}
\item
  \href{http://svmiller.com/blog/2015/12/assorted-tips-students-research-papers/}{Assorted
  Tips for Students on Writing Research Papers}
\item
  \href{https://www.dropbox.com/s/apihjs7di81aqcv/svm-exam-grading-policy.pdf?dl=0}{Exam
  Grading Policy}
\item
  \href{http://svmiller.com/blog/2016/05/fun-with-attendance-grades/}{Fun
  with Attendance and Grades (i.e.~Students Should Attend Class)}
\end{itemize}

\hypertarget{week-02-0121---0125-the-first-topic-where-we-read-john-vasquez}{%
\subsection{Week 02, 01/21 - 01/25: The First Topic Where We Read John
Vasquez}\label{week-02-0121---0125-the-first-topic-where-we-read-john-vasquez}}

\hypertarget{week-03-0128---0201-read-the-nos.-90-97-items-in-my-bib}{%
\subsection{Week 03, 01/28 - 02/01: Read the Nos. 90-97 Items in My
Bib}\label{week-03-0128---0201-read-the-nos.-90-97-items-in-my-bib}}

\emph{Your ``Slow Ride'' appreciation paper is due in Thursday's class}.

\hypertarget{week-04-0204---0208-read-bib-item-no.120}{%
\subsection{Week 04, 02/04 - 02/08: Read Bib Item
No.~120}\label{week-04-0204---0208-read-bib-item-no.120}}

\hypertarget{week-05-0211---0215-the-fourth-topic-with-bib-item-no.510}{%
\subsection{Week 05, 02/11 - 02/15: The Fourth Topic with Bib Item
No.~510}\label{week-05-0211---0215-the-fourth-topic-with-bib-item-no.510}}

\hypertarget{week-06-0218---0222-keep}{%
\subsection{Week 06, 02/18 - 02/22:
Keep}\label{week-06-0218---0222-keep}}

\hypertarget{week-07-0225---0301-going}{%
\subsection{Week 07, 02/25 - 03/01:
Going}\label{week-07-0225---0301-going}}

\hypertarget{week-08-0304---0308-down}{%
\subsection{Week 08, 03/04 - 03/08:
Down}\label{week-08-0304---0308-down}}

\hypertarget{week-09-0311---0315-the}{%
\subsection{Week 09, 03/11 - 03/15: the}\label{week-09-0311---0315-the}}

\hypertarget{week-10-0318---0322-line}{%
\subsection{Week 10, 03/18 - 03/22:
Line}\label{week-10-0318---0322-line}}

\hypertarget{week-11-0325---0329-until}{%
\subsection{Week 11, 03/25 - 03/29:
Until}\label{week-11-0325---0329-until}}

\hypertarget{week-12-0401---0405-you}{%
\subsection{Week 12, 04/01 - 04/05: You}\label{week-12-0401---0405-you}}

\hypertarget{week-13-0408---0412-are}{%
\subsection{Week 13, 04/08 - 04/12: Are}\label{week-13-0408---0412-are}}

\hypertarget{week-14-0415---0419-done}{%
\subsection{Week 14, 04/15 - 04/19:
Done}\label{week-14-0415---0419-done}}

\hypertarget{week-15-0422---0426-with}{%
\subsection{Week 15, 04/22 - 04/26:
with}\label{week-15-0422---0426-with}}

\hypertarget{week-16-0429---0503-your}{%
\subsection{Week 16, 04/29 - 05/03:
your}\label{week-16-0429---0503-your}}

\hypertarget{week-17-0506---0510-syllabus}{%
\subsection{Week 17, 05/06 - 05/10:
Syllabus}\label{week-17-0506---0510-syllabus}}




\end{document}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
    {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}%
}
%\fi
\makeatother