---
title: "Takehome Midterm"
author: "Prof. Field"
date: "3/17/2019"
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

For the takehome midterm you will have 48 hours to complete the objectives listed below. The deadline for submission is 28 March 2019 at 11am EST.

Date/Time started: 17 Mar 2019 10:24am to 12:17, then 12:40 to 2:25
Date/Time completed:

## Loading Libraries

Load all of your libraries in this code block. Indicate why each library is necessary.

```{r Load Libraries, include=FALSE}
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
# cowplot is needed for the cowplot theme for ggplot2
if (!require("ggcorrplot")) install.packages("ggcorrplot"); library(ggcorrplot)
# ggcorrplot is needed for the correlation plot
if (!require("MASS")) install.packages("MASS"); library(MASS)
# MASS is needed for the negative binomial glm function
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
# tidyverse is needed for ggplot2, readr, and dplyr
```

## Objectives for Midterm Exam

* [ ] Import, clean, merge data tables
* [ ] Present graphical summary of Dengue incidence data
* [ ] Data exploration of potential explanatory variables
* [ ] Test Benchmark model of Dengue incidence
* [ ] Improve model of Dengue incidence

## Background

This dataset should be familiar from Lab 6. We will be using the Dengue dataset from a Driven Data competition: 
https://www.drivendata.org/competitions/44/dengai-predicting-disease-spread/

The data for this competition comes from multiple sources aimed at supporting the Predict the Next Pandemic Initiative (https://www.whitehouse.gov/blog/2015/06/05/back-future-using-historical-dengue-data-predict-next-epidemic). 
Dengue surveillance data is provided by the U.S. Centers for Disease Control and prevention, as well as the Department of Defense's Naval Medical Research Unit 6 and the Armed Forces Health Surveillance Center, in collaboration with the Peruvian government and U.S. universities. 
Environmental and climate data is provided by the National Oceanic and Atmospheric Administration (NOAA), an agency of the U.S. Department of Commerce.

The data is provided in two separate files:

1. dengue_features_train: weekly weather and vegetation data for two cities
2. dengue_labels_train: weekly number of dengue cases in each city

There are two cities, San Juan, Puerto Rico and Iquitos, Peru, with test data for each city spanning 5 and 3 years respectively. The data for each city have been concatenated along with a city column indicating the source: *sj* for San Juan and *iq* for Iquitos. 

```{r Read Data}
dengue_features_train <- read_csv("https://s3.amazonaws.com/drivendata/data/44/public/dengue_features_train.csv")
str(dengue_features_train)
summary(dengue_features_train)
dengue_labels_train <- read_csv("https://s3.amazonaws.com/drivendata/data/44/public/dengue_labels_train.csv")
str(dengue_labels_train)
summary(dengue_labels_train)
```

## Feature Descriptions

You are provided the following set of information on a (year, weekofyear) timescale:

(Where appropriate, units are provided as a _unit suffix on the feature name.)

City and date indicators

- city – City abbreviations: sj for San Juan and iq for Iquitos
- week_start_date – Date given in yyyy-mm-dd format

NOAA's GHCN daily climate data weather station measurements

- station_max_temp_c – Maximum temperature
- station_min_temp_c – Minimum temperature
- station_avg_temp_c – Average temperature
- station_precip_mm – Total precipitation
- station_diur_temp_rng_c – Diurnal temperature range

PERSIANN satellite precipitation measurements (0.25x0.25 degree scale)

- precipitation_amt_mm – Total precipitation

NOAA's NCEP Climate Forecast System Reanalysis measurements (0.5x0.5 degree scale)

- reanalysis_sat_precip_amt_mm – Total precipitation
- reanalysis_dew_point_temp_k – Mean dew point temperature
- reanalysis_air_temp_k – Mean air temperature
- reanalysis_relative_humidity_percent – Mean relative humidity
- reanalysis_specific_humidity_g_per_kg – Mean specific humidity
- reanalysis_precip_amt_kg_per_m2 – Total precipitation
- reanalysis_max_air_temp_k – Maximum air temperature
- reanalysis_min_air_temp_k – Minimum air temperature
- reanalysis_avg_temp_k – Average air temperature
- reanalysis_tdtr_k – Diurnal temperature range

Satellite vegetation - Normalized difference vegetation index (NDVI) - NOAA's CDR Normalized Difference Vegetation Index (0.5x0.5 degree scale) measurements

- ndvi_se – Pixel southeast of city centroid
- ndvi_sw – Pixel southwest of city centroid
- ndvi_ne – Pixel northeast of city centroid
- ndvi_nw – Pixel northwest of city centroid

# Takehome Exam Begins here

As a reminder, you may consult your previous homework and group projects, textbook and other readings, and online resources.
Online resources may be used to research ways to solve each problem, but you may not pose questions in online forums about the specific assignment.
You may consult with Prof. Field or with other classmates about technical problems (e.g. where to find a file), but not about how to answer any of the questions.

## Data Wrangling

Use this section to manipulate the two data frames.
1. Follow the Exploratory Data Analysis Checklist (below) to verify the imported data
    a. Ensure that each variable is the appropriate data class and has values that makes sense
    b. For external verification, at a minimum, check that the annual Dengue incidence numbers for each city are realistic
2. Merge the two data frames, verifying that no information was lost during the merge
3. Check the data for NAs both before and after the merge (note that eliminating all rows or columns with NAs will have consequences)

### Check the packaging

```{r}
dim(dengue_features_train)
dim(dengue_labels_train)
colnames(dengue_features_train)
colnames(dengue_labels_train)
sum(dengue_features_train$weekofyear==dengue_labels_train$weekofyear)
sum(dengue_features_train$weekofyear!=dengue_labels_train$weekofyear)
```

### Run str()

```{r}
str(dengue_features_train)
str(dengue_labels_train)
```


### Look at the top and the bottom of your data

```{r}
head(dengue_features_train)
head(dengue_features_train)
head(dengue_labels_train)
tail(dengue_labels_train)
```

### Check your “n”s

```{r}
colSums(is.na(dengue_features_train))
colSums(is.na(dengue_labels_train))
```


### Validate with at least one external data source

```{r}
dengue_labels_train %>%
  group_by(city,year) %>%
  summarise(Weekly_mean=mean(total_cases), Weekly_max=max(total_cases), Yearly_total=sum(total_cases))
```

From: http://www.denguevirusnet.com/epidemiology.html
"The most recent island-wide epidemic occurred in 2007, when more than 10,000 cases were diagnosed." 
In 2007, there were 1969 cases in San Jose- so that makes sense.
"In Peru, nearly 400 cases have been reported until March 2013 of whoch two of them were fatal."
We don't have data for 2013, but the numbers are lower than for Puerto Rico, so that is consistent.

### Merging the data

```{r}
# Make sure that there is a unique index for each table to use for joining
dengue_features_train %>%
  mutate(city_year_week = paste(city, year, weekofyear, sep = "_")) -> dengue_features_train_indexed
dengue_labels_train %>%
  mutate(city_year_week = paste(city, year, weekofyear, sep = "_")) %>% 
  select(-city,-year,-weekofyear) -> dengue_labels_train_indexed
merged_dengue <- merge(dengue_labels_train_indexed, dengue_features_train_indexed, by="city_year_week")
head(merged_dengue)
tail(merged_dengue)
```


### Dealing with the NAs
```{r}
# fill NAs by the latest value
colSums(is.na(merged_dengue))
merged_dengue.filled <- merged_dengue %>% 
  fill(-city_year_week,-total_cases,-city,-year,-weekofyear,-week_start_date)
colSums(is.na(merged_dengue.filled))
```


## What is the average number of cases of Dengue for each week of the year for each city?

Provide a publication-quality graphic to present this comparison. 
You are encouraged to explore options, but only your final graph in this section will be used to evaluate this objective.
Consider the most effective way to illustrate any trends or important comparisons within the data.
Is weekly the most effective bin for incidence? If not, present a better summary.

```{r}
weekly_average <- merged_dengue.filled %>%
  group_by(city,weekofyear) %>%
  summarize(Mean=mean(total_cases), StdDev=sd(total_cases))

weekly_average

# Tested with error bars at first - but this was a bad idea
weekly_average %>%
  ggplot(aes(x=weekofyear, y=Mean, group=city)) + 
  geom_line(aes(color=city)) +
  geom_errorbar(aes(ymin=Mean-StdDev, ymax=Mean+StdDev)) +
  ggtitle("Weekly Dengue Incidence") +
  theme_cowplot()

weekly_average %>%
  ggplot(aes(x=weekofyear, y=Mean, group=city)) + 
  geom_line(aes(color=city)) +
  ggtitle("Weekly Dengue Incidence") +
  xlab("Week of the Year") +
  ylab("Average Dengue Cases") +
  scale_color_discrete(name="City", labels=c("Iquitos","San Juan")) +
  scale_x_continuous(breaks=0:52*4) +
  theme_cowplot()
```


## Data exploration of potential explanatory variables

Consider whether transforming any of the variables might increase the statistical power available.
Explore the correlation of the potential explanatory variables with each other and with dengue incidence.
Present a few publication-quality graphics to illustrate your most important findings.

```{r}
colnames(merged_dengue.filled)
merged_dengue.filled %>%
  dplyr::select(-city_year_week,-city,-year,-weekofyear,-week_start_date) %>%
  summarise_all(.funs= funs(Shapiro=shapiro.test(.)$p.value)) %>%
  t()
## None of them are normal 
merged_dengue.filled %>%
  dplyr::select(-city_year_week,-city,-year,-weekofyear,-week_start_date) %>%
  +0.01 %>%
  log10() %>%
  summarise_all(.funs= funs(Shapiro=shapiro.test(.)$p.value)) %>%
  t()

## log transformation didn't really help with anything
```

```{r}

merged_dengue.filled %>% 
  dplyr::select(-city_year_week,-city,-year,-weekofyear,-week_start_date) %>%
  cor(use = 'pairwise.complete.obs') -> correlation_matrix

ggcorrplot(correlation_matrix, type="lower", hc.order=TRUE, tl.cex=6)

## After some trial and error, I decided to use sqrt of total cases to allow for better visualization (log also worked fine)

ggplot(merged_dengue, aes(x=reanalysis_min_air_temp_k, y=(sqrt(total_cases)))) +
  geom_point() +
  geom_smooth(method='lm')

ggplot(merged_dengue, aes(x=reanalysis_dew_point_temp_k, y=sqrt(total_cases))) +
  geom_point() +
  geom_smooth(method='lm')

ggplot(merged_dengue, aes(x=reanalysis_tdtr_k, y=sqrt(total_cases))) +
  geom_point() +
  geom_smooth(method='lm')

```


## Dengue incidence model

Use a generalized linear model to determine the best model for the weekly incidence of Dengue.
At a first pass consider the "Benchmark" model described here: https://shaulab.github.io/DrivenData/DengAI/Benchmark.html
This model is calculated separately for San Jose and Iquitos and only uses the following variables:
 - reanalysis_specific_humidity_g_per_kg
 - reanalysis_dew_point_temp_k 
 - station_avg_temp_c
 - station_min_temp_c
The code for the Benchmark model uses a machine learning approach to optimize the model.
You should use the model selection approach that we have used in BIOL 364, instead.

```{r}
dengue.glm.1 <- glm.nb(total_cases ~ city + reanalysis_specific_humidity_g_per_kg + reanalysis_dew_point_temp_k + station_avg_temp_c + station_min_temp_c, 
                       data=merged_dengue.filled)

summary(dengue.glm.1)
stepAIC(dengue.glm.1)
```

Graph the model

```{r}
merged_dengue.filled$predicted.1 <- predict.glm(dengue.glm.1, merged_dengue.filled, type="response")
merged_dengue.filled %>% 
  filter(city=="sj") %>%
  mutate(index = as.numeric(row.names(.))) %>%
  ggplot(aes(x = index)) + ggtitle("Benchmark Model for San Jose") +
  geom_line(aes(y = total_cases, colour = "total_cases")) + 
  geom_line(aes(y = predicted.1, colour = "fitted"))

merged_dengue.filled %>% 
  filter(city=="iq") %>%
  mutate(index = as.numeric(row.names(.))) %>%
  ggplot(aes(x = index)) + ggtitle("Benchmark Model for Iquitos") +
  geom_line(aes(y = total_cases, colour = "total_cases")) + 
  geom_line(aes(y = predicted.1, colour = "fitted"))
```

 
## Extend the Benchmark model

Consider and test ways to improve the Benchmark model. 

```{r}
dengue.glm.full <- glm.nb(total_cases ~ city + ndvi_ne + ndvi_nw + ndvi_se + ndvi_sw +
                         precipitation_amt_mm + reanalysis_air_temp_k +
                         reanalysis_avg_temp_k + reanalysis_dew_point_temp_k + reanalysis_max_air_temp_k + reanalysis_min_air_temp_k +
                         reanalysis_precip_amt_kg_per_m2 + reanalysis_relative_humidity_percent + 
                         reanalysis_sat_precip_amt_mm + reanalysis_specific_humidity_g_per_kg + 
                         reanalysis_tdtr_k + station_avg_temp_c + station_diur_temp_rng_c + station_max_temp_c +
                         station_min_temp_c + station_precip_mm, 
                       data=merged_dengue.filled)
summary(dengue.glm.full)
dengue.glm.stepwise <- stepAIC(dengue.glm.full)
summary(dengue.glm.stepwise)
```



```{r}
merged_dengue.filled$predicted.full <- predict.glm(dengue.glm.stepwise, merged_dengue.filled, type="response")
merged_dengue.filled %>% 
  filter(city=="sj") %>%
  mutate(index = as.numeric(row.names(.))) %>%
  ggplot(aes(x = index)) + ggtitle("Full Model for San Jose") +
  geom_line(aes(y = total_cases, color = "total_cases")) + 
  geom_line(aes(y = predicted.full, color = "fitted"))

merged_dengue.filled %>% 
  filter(city=="iq") %>%
  mutate(index = as.numeric(row.names(.))) %>%
  ggplot(aes(x = index)) + ggtitle("Full Model for Iquitos") +
  geom_line(aes(y = total_cases, color = "total_cases")) + 
  geom_line(aes(y = predicted.full, color = "fitted"))
```


### This was my attempt to improve the model with a lag period, but it didn't work very well

Now to repeat but with a 1, 2, 3, or 4 week delay in dengue

```{r}
# First I need to split the data by city
merged_dengue.filled.sj <- filter(merged_dengue.filled, city=="sj")
merged_dengue.filled.iq <- filter(merged_dengue.filled, city=="iq")

merged_dengue.filled.sj$delay1wk <- lag(merged_dengue.filled.sj$total_cases, 1)
merged_dengue.filled.sj$delay2wk <- lag(merged_dengue.filled.sj$total_cases, 2)
merged_dengue.filled.sj$delay3wk <- lag(merged_dengue.filled.sj$total_cases, 3)
merged_dengue.filled.sj$delay4wk <- lag(merged_dengue.filled.sj$total_cases, 4)

merged_dengue.filled.iq$delay1wk <- lag(merged_dengue.filled.iq$total_cases, 1)
merged_dengue.filled.iq$delay2wk <- lag(merged_dengue.filled.iq$total_cases, 2)
merged_dengue.filled.iq$delay3wk <- lag(merged_dengue.filled.iq$total_cases, 3)
merged_dengue.filled.iq$delay4wk <- lag(merged_dengue.filled.iq$total_cases, 4)

merged_dengue.filled.delay <- rbind(merged_dengue.filled.sj, merged_dengue.filled.iq)
```

Now to test the delay models.

```{r}
dengue.glm.full.1wk <- glm.nb(delay1wk ~ city + ndvi_ne + ndvi_nw + ndvi_se + ndvi_sw +
                         precipitation_amt_mm + reanalysis_air_temp_k +
                         reanalysis_avg_temp_k + reanalysis_dew_point_temp_k + reanalysis_max_air_temp_k + reanalysis_min_air_temp_k +
                         reanalysis_precip_amt_kg_per_m2 + reanalysis_relative_humidity_percent + 
                         reanalysis_sat_precip_amt_mm + reanalysis_specific_humidity_g_per_kg + 
                         reanalysis_tdtr_k + station_avg_temp_c + station_diur_temp_rng_c + station_max_temp_c +
                         station_min_temp_c + station_precip_mm, 
                       data=merged_dengue.filled.delay)
summary(dengue.glm.full.1wk)
stepAIC(dengue.glm.full.1wk)
```

```{r}
dengue.glm.full.2wk <- glm.nb(delay2wk ~ city + ndvi_ne + ndvi_nw + ndvi_se + ndvi_sw +
                         precipitation_amt_mm + reanalysis_air_temp_k +
                         reanalysis_avg_temp_k + reanalysis_dew_point_temp_k + reanalysis_max_air_temp_k + reanalysis_min_air_temp_k +
                         reanalysis_precip_amt_kg_per_m2 + reanalysis_relative_humidity_percent + 
                         reanalysis_sat_precip_amt_mm + reanalysis_specific_humidity_g_per_kg + 
                         reanalysis_tdtr_k + station_avg_temp_c + station_diur_temp_rng_c + station_max_temp_c +
                         station_min_temp_c + station_precip_mm, 
                       data=merged_dengue.filled.delay)
summary(dengue.glm.full.2wk)
stepAIC(dengue.glm.full.2wk)
```

```{r}
dengue.glm.full.3wk <- glm.nb(delay3wk ~ city + ndvi_ne + ndvi_nw + ndvi_se + ndvi_sw +
                         precipitation_amt_mm + reanalysis_air_temp_k +
                         reanalysis_avg_temp_k + reanalysis_dew_point_temp_k + reanalysis_max_air_temp_k + reanalysis_min_air_temp_k +
                         reanalysis_precip_amt_kg_per_m2 + reanalysis_relative_humidity_percent + 
                         reanalysis_sat_precip_amt_mm + reanalysis_specific_humidity_g_per_kg + 
                         reanalysis_tdtr_k + station_avg_temp_c + station_diur_temp_rng_c + station_max_temp_c +
                         station_min_temp_c + station_precip_mm, 
                       data=merged_dengue.filled.delay)
summary(dengue.glm.full.3wk)
stepAIC(dengue.glm.full.3wk)
```

```{r}
dengue.glm.full.4wk <- glm.nb(delay4wk ~ city + ndvi_ne + ndvi_nw + ndvi_se + ndvi_sw +
                         precipitation_amt_mm + reanalysis_air_temp_k +
                         reanalysis_avg_temp_k + reanalysis_dew_point_temp_k + reanalysis_max_air_temp_k + reanalysis_min_air_temp_k +
                         reanalysis_precip_amt_kg_per_m2 + reanalysis_relative_humidity_percent + 
                         reanalysis_sat_precip_amt_mm + reanalysis_specific_humidity_g_per_kg + 
                         reanalysis_tdtr_k + station_avg_temp_c + station_diur_temp_rng_c + station_max_temp_c +
                         station_min_temp_c + station_precip_mm, 
                       data=merged_dengue.filled.delay)
summary(dengue.glm.full.4wk)
stepAIC(dengue.glm.full.4wk)
```

## Submitting results of model to Driven Data

```{r}
# submitting the predictions
dengue_features_test <- read_csv("https://s3.amazonaws.com/drivendata/data/44/public/dengue_features_test.csv")

# checking that the format matches
head(dengue_features_test)

# dealing with the NAs
colSums(is.na(dengue_features_test))
dengue_features_test.filled <- dengue_features_test %>% 
  fill(-city,-year,-weekofyear,-week_start_date)
colSums(is.na(dengue_features_test.filled))

dengue_features_test.filled$predicted <- predict.glm(dengue.glm.stepwise, dengue_features_test.filled, type="response")

submissions = read.csv("https://s3.amazonaws.com/drivendata/data/44/public/submission_format.csv")
inner_join(submissions, rbind(dengue_features_test.filled)) %>%
  dplyr::select(city, year, weekofyear, total_cases = predicted) ->
predictions

predictions$total_cases %>% round() -> predictions$total_cases
write.csv(predictions, 'predictions1.csv', row.names = FALSE)
```


# Acknowledgements

Correlation plot adapted from:
How to use predict():
How to format to submit to Driven Data:
https://shaulab.github.io/DrivenData/DengAI/Benchmark.html
