# R code to generate the linear regression model

Here is my regression model:

```{r model, collapse=TRUE}
bradford <- values[["MyDataFrame"]]
options(digits = 4)
fit <- lm(Absorbance ~ Concentration, data = bradford)
modelCoefficients <- coef(fit)
summary(fit)
```

The fitting result is $Absorbance =  `r modelCoefficients[2]`Concentration + `r modelCoefficients[1]`$

The adjusted R squared is
```{r R squared, collapse=TRUE}
summary(fit)$adj.r.squared
```

This linear regression model tested the hypothesis that there is a linear relationship between *Aborbance* and *Concentration*. 
The probability that this relationship is more likely than the null hypothesis (that *Aborbance* does NOT depend on *Concentration*):

```{r p value, collapse=TRUE}
summary(fit)$coefficients[,4][2] 
```


Below is a scatter plot with the regression line.

```{r plot, fig.height=5}
bradford <- values[["MyDataFrame"]]
plot(Absorbance ~ Concentration, data = bradford, pch = 19,
     xlab = "Protein concentration (mg/mL)", 
     ylab = "Absorbance (at 595 nm)")
abline(lm(Absorbance ~ Concentration, data = bradford), lwd = 2)
```
