---
title: "Lab 04B"
author: "Biology Student"
date: "2/3/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries
Use the following syntax to load any libraries that you need. This code will prevent errors that will interfere with knitting of the Rmd file.
Also note that the `include=FALSE` option will prevent this code chunk from appearing in the markdown file.

We will load the entire tidyverse library, which includes *dplyr*, *ggplot2*, *readr*, and other packages.

```{r Load Libraries, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
```

## Loading data

For this exercise we will be using data from Prof. Reeder's fieldwork in Uganda. Epauletted fruit bats were collected at two different
times of the year for this study and a variety of health metrics were measured.

Note that this data should not be shared outside this class without permission.

```{r}
BatData <- read_csv("UgandaBatsFilteredMetrics.csv", 
    col_types = cols(`Collection date` = col_date(format = "%m/%d/%y")))
# The col_date format was used to make sure that the Collection date was interpretted properly.
# See ?as.Date for more information and examples.

# Below is a very handy way to eliminate those annoying spaces in the column names!
names(BatData) <- make.names(names(BatData))

```


