---
title: "Homework 03"
author: "Biology Student"
date: "2/3/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries
Use the following syntax to load any libraries that you need. This code will prevent errors that will interfere with knitting of the Rmd file.
Also note that the `include=FALSE` option will prevent this code chunk from appearing in the markdown file.

We will load the entire tidyverse library, which includes *dplyr*, *ggplot2*, *readr*, and other packages.

```{r Load Libraries, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
```

# Exploratory Data Analysis Checklist

Follow Chapter 5 from Exploratory Data Analysis with R by Roger D. Peng
Explore the Uganda Bat Dataset, formulating a question and exploring that question.

## Formulate your question

## Read in your data

## Check the packaging

## Run str()

## Look at the top and the bottom of your data

## Check your “n”s

## Validate with at least one external data source

## Try the easy solution first

## Challenge your solution

## Follow up

# Analytic Graphics

Follow Section 6.1 and 6.2 to visualize the comparisons that were essential to answering the question.

## Show comparisons

## Show causality, mechanism, explanation, systematic structure
