---
title: "Traditional Research Practices"
author: "Ken Field"
date: "February 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in survey results

Students in Biology 364/664 Analysis and Visualization of Biological Data and 
faculty in the Biology department at Bucknell separately took a survey about "Traditional Research Practices."

The survey was identical to the one used for Fraser, et al., Questionable research practices in ecology and evolution,
PLOS One, https://doi.org/10.1371/journal.pone.0200303


```{r Read Results and convert to factors}
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
Faculty <- read_csv("../04-Data/Traditional Research Practices Survey (Responses) - Form Responses 1.csv", 
    col_types = cols(Timestamp = col_skip()))
Students <- read_csv("../04-Data/Traditional Research Practices Survey 2 (Responses) - Form Responses 1.csv", 
    col_types = cols(Timestamp = col_skip()))
str(Faculty)
str(Students)
Faculty <- mutate_if(Faculty, is.character, factor)
Students <- mutate_if(Students, is.character, factor)
str(Faculty)
str(Students)
```

## Plotting results


```{r Bargraphs}
ggplot(mpg, aes(class)) + 
  geom_bar()
ggplot(Faculty, aes(`Reporting a set of statistical models as the complete tested set when other candidate models were also tested.`)) + 
    geom_bar(aes(fill = `Reporting a set of statistical models as the complete tested set when other candidate models were also tested.`)) + coord_flip()
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
