---
title: "Homework 02"
output: github_document
author: "Biol 364 Student"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## The NCHA-II Dataset
Follow this link to see copies of the Web surveys completed by Bucknell students in 2014 and 2016.
Note that there are some differences in the surveys in the two years.
https://www.acha.org/NCHA/About_ACHA_NCHA/Survey/NCHA/About/Survey.aspx

```{r Importing Data}
library(haven)
NCHA <- read_sav("NCHA-II_WEB_SPRING_2014_and_2016_for_Ken_Field_class_project.SAV")
View(NCHA)
```

Each of the columns has a name that refers to the question number on the web-based form. It also has a label that gives a short version of the question.
However, R doesn't know how to handle this data type very well, so you will often want to tell R to convert the column into a factor or a numeric type.
We also want to convert the GPA question into a format that makes more sense. We will make a new column "GPA" that converts the survey results into a 4 point scale.

63) What is your approximate cumulative grade average?
 A
 B
 C
 D/F
 N/A

```{r GPA}
library(UsingR)
summary(NCHA$NQ63)
#simple.eda(NCHA$NQ63) ### This will give an error because of the data type
simple.eda(as.numeric(NCHA$NQ63))
NCHA$GPA <- 5 - as.numeric(NCHA$NQ63)
```



```{r Data Exploration}
dim(NCHA)
summary(NCHA)
summary(NCHA$NQ43)
simple.eda(as.numeric(NCHA$NQ43))
```

After exploring the dataset, formulate some tentative hypotheses that you wish to test.

```{r Data Visualization}
library(ggplot2)
library(cowplot)
ggplot(NCHA) +
  aes(x = NQ43) +
  geom_bar() +
  theme_cowplot()

mosaicplot(~ GPA + NQ43, data = NCHA, color = TRUE)

ggplot(NCHA) +
  aes(x = as.numeric(NQ43), y = GPA) +
  geom_jitter(aes(color = factor(NQ51)), na.rm=TRUE) +
  theme_cowplot() +
  theme(legend.position = "right") +
  labs(color="Class Year") + 
  xlab("Problem with sleepiness (last 7 days)") + 
  ylab("Approximate GPA")
```

43) People sometimes feel sleepy during the daytime. In the past 7 days, how much of a
problem have you had with sleepiness (feeling sleepy, struggling to stay awake) during your
daytime activities?
 No problem at all
 A little problem
 More than a little problem
 A big problem
 A very big problem

51) What is your year in school?
 1st year undergraduate
 2nd year undergraduate
 3rd year undergraduate
 4th year undergraduate
 5th year or more undergraduate
 Graduate or professional
 Not seeking a degree
 Other

```{r Statistical Tests}

```

Acknowledgements
=========================

